---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r loadactivity}
library(reshape2)
library(lubridate)
activity <- read.csv("activity.csv", header = T)
```

## Let's melt the data so that we can calculate total steps per day and average steps across intervals later
```{r }
melted_days <- melt(activity, id.vars = "date", value.name = "steps")
melted_intervals <- melt(activity, id.vars = "interval", measure.vars="steps", value.name = "steps", na.rm = T)
```
## What is mean total number of steps taken per day?
```{r extracting total activity per day, echo=FALSE}
#Let's calculate the total steps per day 
total_steps <- dcast(melted_days, date~variable, sum)

#Let's draw a histogram of the steps. Breaks = 25, so that each bar represents a 1000
hist(total_steps$steps, xlab = "Total Activity per day", breaks=25 )
```

##Lets calculate the mean and median for the total steps
```{r meanTotalSteps, echo=F}
meanTotalSteps = mean(total_steps$steps,na.rm = T)
medianTotalSteps = median(total_steps$steps, na.rm=T)
```

## What is the average daily activity pattern?
```{r calculating average steps per day}
average_daily_activity <- dcast(melted_intervals,interval~variable, mean)

#Lets plot this so that we can visualize what is the most prominent interval for activities
plot(type="l",y=average_daily_activity$steps, x=average_daily_activity$interval, xlab="5 min interval", ylab="Average Steps Taken")
```

##Let's find the interval with the max average daily value
```{r }
average_daily_activity[which.max(average_daily_activity$steps),]$interval

```


## Imputing missing values
###Let's calculate the total number of rows with missing values
length(which(!complete.cases(activity)))

###Now we will use a simple recoding. We will take the mean total steps and divide it by 288 and replace the NA with this value
meanIntervalSteps <- meanTotalSteps/288
activity_recoded <- activity
activity_recoded$steps[is.na(activity_recoded$steps)] <- meanIntervalSteps

###Now we will calculate the total steps and the mean steps using these values
```{r }
melted_days_recoded <- melt(activity_recoded, id.vars = "date", value.name = "steps")
melted_intervals_recoded <- melt(activity_recoded, id.vars = "interval", measure.vars="steps", value.name = "steps", na.rm = T)
```
```{r extracting total activity per day with recoded values, echo=FALSE}
#Let's calculate the total steps per day 
total_steps_recoded <- dcast(melted_days_recoded, date~variable, sum)

#Let's draw a histogram of the steps. Breaks = 25, so that each bar represents a 1000
hist(total_steps_recoded$steps, xlab = "Total Activity per day", breaks=25 )
```

```{r meanTotalStepsRecoded, echo=F}
meanTotalStepsRecoded = mean(total_steps_recoded$steps,na.rm = T)
medianTotalStepsRecoded = median(total_steps_recoded$steps, na.rm=T)
```
### We see that there are no differences in the mean and median values when you recode missing values
### However the distribution now has a lot of values at the mean and looks strongly normalized


## Are there differences in activity patterns between weekdays and weekends?